---
title: "RLisbonaMSDS6306_Week4_BootstrapSampling"
author: "Randy Lisbona"
date: "June 4, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
require("boot")
require("psych")
knitr::opts_chunk$set(echo = TRUE)
```

## Create the normal and exponential sample datasets

Use rnorm and rexp to create some sample datasets

```{r Samplesets,echo = FALSE}
Normal.dataset.n30mu10s1 <- rnorm(30,10,1)
Normal.dataset.n10mu10s1 <- rnorm(10,10,1)
Normal.dataset.n30mu15s1 <- rnorm(30,15,1)

Exponential.dataset.n30lambda1.0 <- rexp(30,1)
Exponential.dataset.n10lambda1.0 <- rexp(10,1)
#Exponential.dataset.n30lambda0.25 <- rexp(30,.25)
```
## Print the datasets, inlcude the first 10 records from each dataset
```{r printheader,echo = FALSE}
head(Normal.dataset.n30mu10s1, n= 10)
head(Normal.dataset.n10mu10s1, n= 10) 
#head(Normal.dataset.n30mu15s1, n= 10) 

head(Exponential.dataset.n30lambda1.0, n= 10) 
head(Exponential.dataset.n10lambda1.0, n= 10)
#head(Exponential.dataset.n30lambda0.5, n= 10) 

```

## Explore the data with Plot of the normal and exponential sample datasets

Use plot and hist to compare the datasets
set x and y limits to make it easier to compare plots

```{r plotsunsorted,echo = FALSE}

par(mfrow=c(4,2),mar=c(1,1,1,1))

plot(Normal.dataset.n30mu10s1,xlim=c(0,30), ylim=c(0,30), main="Normal Dist, n=30, mu=10, sd=1")
plot(Exponential.dataset.n30lambda1.0,xlim=c(0,30), ylim=c(0,30), main="Exp Dist, n=30, lambda = 1")

plot(Normal.dataset.n10mu10s1,xlim=c(0,30), ylim=c(0,30), main="Normal Dist, n=10, mu=10, sd=1")
plot(Exponential.dataset.n10lambda1.0,xlim=c(0,30), ylim=c(0,30), main="Exp Dist, n=10, lambda = 1")

#plot(Normal.dataset.n30mu15s1,xlim=c(0,30), ylim=c(0,30), main="Normal Dist, n=30, mu=15, sd=1")
#plot(Exponential.dataset.n30lambda0.25,xlim=c(0,30), ylim=c(0,30), main="Exp Dist, n=30, lambda = 0.25")
```

## Sort the records ascending and plot again. Include histograms
set x and y limits to make it easier to compare plots

```{r sorted,echo = FALSE}

par(mfrow=c(4,2),mar=c(1,1,1,1))

plot(sort(Normal.dataset.n30mu10s1), xlim=c(0,30), ylim=c(0,30), main="Normal Dist, n=30, mu=10, sd=1")
hist(Normal.dataset.n30mu10s1,breaks = 10,xlim=c(0,30), main="Normal Dist, n=30, mu=10, sd=1")

plot(sort(Normal.dataset.n10mu10s1), ylim=c(0,30), main="Normal Dist, n=10, mu=10, sd=1")
hist(Normal.dataset.n10mu10s1,breaks = 10,xlim=c(0,30), main="Normal Dist, n=10, mu=10, sd=1")

#plot(sort(Normal.dataset.n30mu15s1), ylim=c(0,30), main="Normal Dist, n=30, mu=15, sd=1")
#hist(Normal.dataset.n30mu15s1,breaks = 10,xlim=c(0,30), main="Normal Dist, n=30, mu=15, sd=1")

plot(sort(Exponential.dataset.n30lambda1.0), ylim=c(0,30), main="Exp Dist, n=30, lambda = 1")
hist(Exponential.dataset.n30lambda1.0,breaks = 10,xlim=c(0,30), ylim=c(0,10), main="Exp Dist, n=30, lambda = 1")

plot(sort(Exponential.dataset.n10lambda1.0), ylim=c(0,30), main="Exp Dist, n=10, lambda = 1.0")
hist(Exponential.dataset.n10lambda1.0,breaks = 10, xlim=c(0,30),ylim=c(0,10), main="Exp Dist, n=10, lambda = 1")

#plot(sort(Exponential.dataset.n30lambda0.25), ylim=c(0,30), main="Exp Dist, n=30, lambda = 0.25")
#hist(Exponential.dataset.n30lambda0.25,breaks=10, xlim=c(0,30),ylim=c(0,10), main="Exp Dist, n=30, lambda = 0.25")

```
## Resample the datasets, compare original histogram to resampled histogram
# notice that the resampled histograms closely resemble a normal distribution, illustrating the central limit theorem

```{r resample ,echo = FALSE}
R <- 1000
Normal.dataset.n30mu10s1.resample <- numeric(R)
Normal.dataset.n10mu10s1.resample <- numeric(R)
#Normal.dataset.n30mu15s1.resample <- numeric(R)
Exponential.dataset.n30lambda1.0.resample <- numeric(R)
Exponential.dataset.n10lambda1.0.resample <- numeric(R)
#Exponential.dataset.n30lambda0.25.resample <- numeric(R)

############################################################
x=Normal.dataset.n30mu10s1
mean(x)
for (i in 1:R) {
  sample <- sample(x, size = length(x), replace = TRUE)
  Normal.dataset.n30mu10s1.resample[i] = mean(sample)
}

par(mfrow = c(1,2))
hist(Normal.dataset.n30mu10s1, main="Original dataset", xlab="Normal Dist, n=30, mu=10, sd=1")
hist(Normal.dataset.n30mu10s1.resample, main="Bootstrap resampled", xlab="Normal Dist")
print("Original dataset")
describe(Normal.dataset.n30mu10s1)
print("Resampled dataset")
describe(Normal.dataset.n30mu10s1.resample)

###########################################################
x=Normal.dataset.n10mu10s1
mean(x)
for (i in 1:R) {
  sample <- sample(x, size = length(x), replace = TRUE)
  Normal.dataset.n10mu10s1.resample[i] = mean(sample)
}

par(mfrow = c(1,2))
hist(Normal.dataset.n10mu10s1, main="Original dataset", xlab="Normal Dist, n=10, mu=10, sd=1")
hist(Normal.dataset.n10mu10s1.resample, main="Bootstrap resampled", xlab="Normal Dist")
print("Original dataset")
describe(Normal.dataset.n10mu10s1)
print("Resampled dataset")
describe(Normal.dataset.n10mu10s1.resample)
################################################################
# x=Normal.dataset.n30mu15s1
# mean(x)
# for (i in 1:R) {
#   sample <- sample(x, size = length(x), replace = TRUE)
#   Normal.dataset.n30mu15s1.resample[i] = mean(sample)
# }
# 
# par(mfrow = c(1,2))
# hist(Normal.dataset.n30mu10s1, main="Original dataset", xlab ="Normal Dist, n=30, mu=15, sd=1")
# hist(Normal.dataset.n30mu10s1.resample, main="Bootstrap resampled", xlab ="Normal Dist, n=30, mu=15, sd=1")
# describe(Normal.dataset.n30mu15s1)
# describe(Normal.dataset.n30mu15s1.resample)
# 


##############################################################
##########################################################
x=Exponential.dataset.n30lambda1.0
resamplex = Exponential.dataset.n30lambda1.0.resample
mean(x)
for (i in 1:R) {
  sample <- sample(x, size = length(x), replace = TRUE)
  resamplex[i] = mean(sample)
}

par(mfrow = c(1,2))
hist(x, main="Original dataset", xlab="Exponential, n=30, lambda = 1.0")
hist(resamplex, main="Bootstrap resampled", xlab="Exponential")
print("Original dataset")
describe(x)
print("Resampled dataset")
describe(resamplex)

##########################################################
x=Exponential.dataset.n10lambda1.0
resamplex = Exponential.dataset.n10lambda1.0.resample
mean(x)
for (i in 1:R) {
  sample <- sample(x, size = length(x), replace = TRUE)
  resamplex[i] = mean(sample)
}

par(mfrow = c(1,2))
hist(x, main="Original dataset", xlab="Exponential, n=10, lambda = 1.0")
hist(resamplex, main="Bootstrap resampled", xlab="Exponential")
print("Original dataset")
describe(x)
print("Resampled dataset")
describe(resamplex)



```

## Conclusion
# The bootstrap method can be used to create a sample distribution from small data sets that approximates a normal sample from the original population , demonstrating the central limit theorem.
